---
title: "finalproject_dnkarp"
author: "dnkarp"
date: "11/02/2018"
output: html_document
---

#--------------------------------------------------------------
"Social impact could include, but not limited to; workforce education/development, workforce diversity, neighborhood revitalization, health and environment friendly and job creation." - WNY Impact Investment Fund

RULE: a low-income community census tract has an individual poverty rate of at least 20% and median family income up to 80% percent of the area median. Every state or territory can designate up to 25% of its census tracts that meet qualification requirements

EXCEPTION: 5% of census tracts that do not meet the definition of a low-income community can be designated under an exemption. Exempt census tracts must be contiguous with low-income community census tracts that are designated as Opportunity Zones, and the median family income of the exempt tract must not exceed 125% of the median family income of the designated low-income community census tract with which it is contiguous. 

DEFINITIONS: low-income community 45d(e) [American Jobs Creation Act of 2004, IRC §45D(e)(2)]
A “low-income community” is defined as any population census tract where the 
- poverty rate for such tract is at least 20% 
- or, in the case of a tract not located within a metropolitan area, median family income for such tract does not exceed 80 of statewide median family income, 
- or, in the case of a tract located within a metropolitan area, the median family income for such tract does not exceed 80% of the greater of statewide median family income or the metropolitan area median family income.

--------------------------------------------------------------
exploring social impact investments -- how many CDFIs, CDEs, CDCs are opporating in the 'region'... NMTCs?
number of impact funds?
indicators of:
  workforce education/development, 
  workforce diversity, 
  neighborhood revitalization, 
  health and environment friendly,
  and job creation.
  
bond issuance?

#--------------------------------------------------------------
### 0.Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
#--------------------------------------------------------------
### 1.Load libraries
```{r load libs, message=F, warning=F, cache=T, results='hide'}

#install.packages(c("tidyverse", "broom", "data.table", "purrr", "stringr", "readxl", "skimr", "units", "sf", "maps", "albersusa", "tidycensus", "tigris", "viridis", "gridExtra", "ggthemes", "cowplot", "tableone", "ggpubr"))

# data wrangling libs
library(tidyverse)
library(broom)
library(data.table)
library(purrr)
library(stringr)
library(readxl)

# data analysis
library(skimr)
library(units)
library(tableone)

# spatial libs
library(sf)
library(maps)
library(albersusa)

# census data libs + install api key
library(tidycensus)
library(tigris)
options(tigris_use_cache = TRUE)

# graphics/viz libs
library(viridis)
library(gridExtra)
library(ggthemes)
library(cowplot)
library(ggpubr)


```

### 2.Load data
##### 2a.Load list of designated 'Opportunity Zones' from csv file
```{r 2a download + load csv data, message=F, warning=F, cache=T}

# A list of US census tracts that are designated "Opportunity Zones" is available as csv(source: CDFI Fund[https://www.cdfifund.gov/Pages/Opportunity-Zones.aspx]); csv, tract-level data, entire US, n = 74134

# check is csv exists in data folder; if not, download and extract
if(!file.exists("data/Opportunity Zones.csv")){
# set URL of data source
data.url <- "https://www.cdfifund.gov/Documents/Opportunity%20Zones2.zip"
# set file path
data.file <- 'data/ozone_data_from_cdfifund.zip'
# download data to local directory
download.file(data.url,data.file, method='curl')
# unzip archived file
unzip(data.file, exdir = "data", files = c("Opportunity Zones.csv"), overwrite = TRUE)
}

# load csv into data frame
ozones.raw <- read_csv("data/Opportunity Zones.csv")

```

##### 2b.Clean 'Opportunity Zones' (OZone) data frame
```{r 2b load csv data, message=F, warning=F, cache=T}

# get list of state postal codes (2-letter abbreviations), assign a state/territory flag
# create a logical vector, true for 50 states +DC; false for territories (PR - Puerto Rico, VI - US Vir Islands, GU - Guam, MP - N.Mariana, UM - Minor Outlying, AS - Am Samoa)
stab <- rbind(
   cbind(state=c(unique(fips_codes$state)[1:51]),is.state=as.logical(T))
  ,cbind(state=c(unique(fips_codes$state)[52:57]),is.state=as.logical(F))
)

# add state letter codes, and state/territory flag to ozone tract-level data;  
ozones <- ozones.raw %>% 
  mutate("state_code"=substr(ozones.raw$GEOID10,1,2)) %>% #extract 2-digit state FIPS codes from 11-digit tract FIPS
  merge(data.frame(unique(fips_codes[,1:2])),by="state_code") %>% #merge (left join) state letter codes based on FIPS codes
  merge(stab,by="state") #merge state/territory logic flag

# set ozone datsframe as datatable
setDT(ozones)

# subset ozone data into 50+DC and territories
ozones.51 <- ozones[as.logical(is.state),]
ozones.terr <- ozones[!as.logical(is.state),]

```

##### 2c.Load spatial features data from ACS
```{r 2c download + load ACS spatial features data, cache=T}

# TRACT-level data
# check is RDS file exists in data folder; if not, download
if(!file.exists("data/ustracts_sf.rds")){
# download census tract boundaries (with population data, 2016) using tidycensus; get per state and bind
ustracts_sf <- reduce( #can only request tracts per state, so it's necessary to loop through each state
  purrr::map(stab[stab[,2]==T,1], function(x) { #loop through statelist, for 'states' only (exclude territories) 
    get_acs(geography = "tract", variables = "B01003_001", #include population variable -- `get_acs` requires at least 1 var
            state = x, geometry = TRUE) # include geometries
  }), 
  rbind #bind (stack rows) into complete data frame
)
# Save a sf object to a file
saveRDS(ustracts_sf, "data/ustracts_sf.rds")
}

# check is object exists in environment; if not, reload
if(!exists("ustracts_sf")){
ustracts_sf <- readRDS("data/ustracts_sf.rds")
}

# COUNTY-level data
# check is RDS file exists in data folder; if not, download
if(!file.exists("data/uscounties_sf.rds")){
# get counties (filter out Puerto Rico)
uscounties_sf <- get_acs(geography = "county", variables = "B01003_001", geometry = TRUE) %>% filter(substr(GEOID,1,2)!="72")
# Save a sf object to a file
saveRDS(uscounties_sf, "data/uscounties_sf.rds")
}

# check is object exists in environment; if not, reload
if(!exists("uscounties_sf")){
uscounties_sf <- readRDS("data/uscounties_sf.rds")
}

# STATE-level data
# check is RDS file exists in data folder; if not, download
if(!file.exists("data/usstates_sf.rds")){
# get states (filter Puerto Rico)
usstates_sf <- get_acs(geography = "state", variables = "B01003_001", geometry = TRUE) %>% filter(GEOID!="72")
# Save a sf object to a file
saveRDS(usstates_sf, "data/usstates_sf.rds")
}

# check is object exists in environment; if not, reload
if(!exists("usstates_sf")){
usstates_sf <- readRDS("data/usstates_sf.rds")
}

```

##### 2d.Clean spatial features data
```{r 2d load ACS spatial features data from repo, cache=T}

# reproject to albers & drop hawaii and alaska (for now; look into moving HI & AK for single layout)
lower48 <- data.frame(state_code=unique(fips_codes$state_code)) %>% #get vector of FIPS codes 
  filter(!(state_code %in% c("02","15","60", "66", "69", "72", "74", "78"))) %>% #exclude territories and HI/AK
  mutate(state_code=as.character(state_code)) %>% #set numeric FIPS code as character
  .[[1]] #convert dataframe to character vector 

# define function to filter by lower48 & reproject to albers
prj.alb.48 <- function(x){x %>% 
    filter(substr(GEOID,1,2) %in% lower48) %>% 
    st_transform(us_laea_proj)}

# filter/repoject tracts
ustracts_sf.48.albers <- prj.alb.48(ustracts_sf)
# filter/repoject counties
uscounties_sf.48.albers <- prj.alb.48(uscounties_sf)
# filter/repoject states
usstates_sf.48.albers <- prj.alb.48(usstates_sf)

```

##### 2e.Load New Market Tax Credit (NMTC) data
```{r 2e download + load NMTC data from xlsx file, cache=T}

# A list of NMTC investment projects, including date, location, project info; source: CDFI Fund, NMTC Public Data Release: 2003-2015 Data File [https://www.cdfifund.gov/news-events/news/Pages/news-detail.aspx?NewsID=270&Category=Press%20Releases]

# check is csv exists in data folder; if not, download and load
if(!file.exists("data/FY 2017 NMTC Public Data Release_v2.xlsx")){
# set URL of data source
data.url <- 'https://www.cdfifund.gov/Documents/FY%202017%20NMTC%20Public%20Data%20Release_v2.xlsx'
# set file path
data.file <- 'data/FY 2017 NMTC Public Data Release_v2.xlsx'
# download data to local directory
download.file(data.url,data.file, method='curl')
}

# import from excel doc
nmtc.raw <- read_excel('data/FY 2017 NMTC Public Data Release_v2.xlsx', sheet = 3)

# clean & format data
nmtc <- nmtc.raw %>% 
  mutate(GEOID = str_pad(`2010 Census Tract`, width=11, side="left", pad="0")) %>% 
  mutate(metro = as.factor(`Metro/Non-Metro, 2010 Census`)) %>% 
  mutate(cde = as.factor(`Community Development Entity (CDE) Name`)) %>% 
  mutate(type = as.factor(`QALICB Type`)) %>% 
  mutate(year = `Origination Year`) %>% 
  mutate(qlici.amount = `Project QLICI Amount`) %>% 
  mutate(total.cost = `Estimated Total Project Cost`) %>% 
  select(GEOID, metro, cde, type, year, qlici.amount, total.cost)

# summarize data from project level to census tract, create summary variables
nmtc.tract <- nmtc %>%
  group_by(GEOID, type) %>% 
  mutate(one = 1) %>% 
  summarize(one = sum(one)) %>% 
  spread(type, one) %>% 
  group_by(GEOID) %>% 
  summarise_all(funs(sum))

nmtc.tract <- nmtc %>% 
  group_by(GEOID) %>% 
  summarize(metro = first(metro), 
            n.cde = length(unique(cde)), 
            n.prj = n(), 
            avg.prj.yr = n()/length(unique(year)), 
            ttl.amt.prj = sum(qlici.amount), 
            avg.amt.prj = mean(qlici.amount), 
            avg.amt.yr = sum(qlici.amount)/length(unique(year))) %>%
  merge(nmtc.tract, by="GEOID")

```
##### 2f.Load regional characteristics (ACS variables)
```{r 2f download + load attribute data from ACS, cache=T}

# load variables (2016 5-year estimates)
v16 <- load_variables(2016, "acs5", cache = TRUE)
if(F){ # this can be used as a look up
  v16 %>% select(name, label) %>% filter(grepl("19113", name))
  v16 %>% select(name, label) %>% filter(grepl("Poverty", label))
}

## DEFINE VARIABLE SELECTION ##

# Population Variables
# TotalPopulation[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&var=B01003001]
totpop_var <- cbind(c("B01003_001"),c("TotPop"))

# Income Variables
# BelowPoverty[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&var=B17001002]
# MedianFamilyIncome[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&var=B19113001]
povinc_var <- cbind(
  c("B17001_001","B17001_002","B19113_001"),
  c("PovPopDenominator","BelowPov","MdnFamInc")
  )

# Race/Ethnicity Variables
# Race[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001]
race_var <- cbind(
  c("B02001_002","B02001_003","B02001_004","B02001_005","B02001_006"),
  c("White","Black","NatAm","Asian","NatIsl")
  )

# Socioeconomic Variables 
# Education[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B15003]
# Employment[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B23025]
# Housing1[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B25002]
# Housing2[https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B25003]
socioecon_var <- cbind(
  c("B15003_001", "B15003_017", "B15003_018", "B15003_019","B15003_020", "B15003_021","B15003_022","B15003_023","B15003_024", "B15003_025", "B23025_003", "B23025_005", "B25002_001", "B25002_003", "B25003_001", "B25003_003"),
  c("Pop25", "HS", "GED", "CollegeNoDegree0", "CollegeNoDegree1", "AssocDegree", "BachDegree", "MastDegree", "ProfDegree", "DoctDegree", "CivLabor", "Unempl", "HousingUnits", "Vacant", "OccUnits", "OccRental" )
  )

# Zip code and County level data
# commerical development, jobs, industry, GDP??
# urban: urban/suburban/rural/metro-region
# health: diabetes, obesity, substance abuse, chronic
# buit environment: parks, rec centers, health clinics, civic inst?
# real estate: property value? (Zillow)

# merge variable name vectors into dataframe
select_vars <- rbind(totpop_var, povinc_var, race_var, socioecon_var) %>% 
  data.frame() %>% 
  rename(variable = X1, v = X2) %>% 
  mutate(variable = as.character(variable))

# get statelist vector 
statelist <- stab[stab[,2]==T,1]

## GET DATA ##

# Get TRACT-level data
if(!file.exists("data/ustrctsacs16_df.rds")){
# get 2016 data
us.trcts.acs16 <- reduce(
  purrr::map(statelist, function(x) { # loop through statelist
    get_acs(geography = "tract", year = 2016, variables = select_vars[,1], # get tract level data for 2016
            state = x, geometry = F) # pull for each state, x, do not include geometry
  }), 
  rbind # bind tracts for each state into one data frame
)
# Save a sf object to a file
saveRDS(us.trcts.acs16, "data/ustrctsacs16_df.rds")
}

# check is object exists in environment; if not, reload
if(!exists("us.trcts.acs16")){
us.trcts.acs16 <- readRDS("data/ustrctsacs16_df.rds")
}

if(!file.exists("data/ustrctsacs15_df.rds")){
# get 2015 data (2010 data is available, but returns an error)
us.trcts.acs15 <- reduce(
  purrr::map(statelist, function(x) {
    get_acs(geography = "tract", year = 2015, variables = select_vars[,1], 
            state = x, geometry = F)
  }), 
  rbind
)
# Save a sf object to a file
saveRDS(us.trcts.acs15, "data/ustrctsacs15_df.rds")
}

# check is object exists in environment; if not, reload
if(!exists("us.trcts.acs15")){
us.trcts.acs15 <- readRDS("data/ustrctsacs15_df.rds")
}

# Get COUNTY-level data
if(!file.exists("data/uscntyacs16_df.rds")){
# get 2016 data
us.cnty.acs16 <- get_acs(geography = "county", year = 2016, variables = select_vars[,1],geometry = F) 
# Save a sf object to a file
saveRDS(us.cnty.acs16, "data/uscntyacs16_df.rds")
}

# check is object exists in environment; if not, reload
if(!exists("us.cnty.acs16")){
us.cnty.acs16 <- readRDS("data/uscntyacs16_df.rds")
}

# Get STATE-level data
if(!file.exists("data/usstacs16_df.rds")){
# get 2016 data
us.st.acs16 <- get_acs(geography = "state", year = 2016, variables = select_vars[,1],geometry = F) 
# Save a sf object to a file
saveRDS(us.st.acs16, "data/usstacs16_df.rds")
}

# check is object exists in environment; if not, reload
if(!exists("us.st.acs16")){
us.st.acs16 <- readRDS("data/usstacs16_df.rds")
}

## FORMAT DATA ##

# transform long data to wide format
#TRACT 2016
us.trcts.acs16.wide <- us.trcts.acs16 %>% 
  left_join(select_vars, by="variable") %>% # join variable names to varible codes
  select(-c(moe,NAME,variable)) %>% # drop unnecessary vars
  spread(key = v, value = estimate, sep = "16.") #use `spread` to generate collumns; add v16. for each variable (year 2016)
#TRACT 2015
us.trcts.acs15.wide <- us.trcts.acs15 %>% 
  left_join(select_vars, by="variable") %>% 
  select(-c(moe,NAME,variable)) %>% 
  spread(key = v, value = estimate, sep = "15.") #add v15. for each variable (year 2015)
#COUNTY
us.cnty.acs16.wide <- us.cnty.acs16 %>% 
  left_join(select_vars, by="variable") %>% # join variable names to varible codes
  select(-c(moe,NAME,variable)) %>% # drop unnecessary vars
  spread(key = v, value = estimate, sep = "16.") #use `spread` to generate collumns; add v16. for each variable (year 2016)
#STATE
us.st.acs16.wide <- us.st.acs16 %>% 
  left_join(select_vars, by="variable") %>% # join variable names to varible codes
  select(-c(moe,NAME,variable)) %>% # drop unnecessary vars
  spread(key = v, value = estimate, sep = "16.") #use `spread` to generate collumns; add v16. for each variable (year 2016)

#----------------------------------
# merge (left join) 2015 to 2016 tracts
us.trcts.acs.wide <- merge(us.trcts.acs16.wide,us.trcts.acs15.wide,by="GEOID")

# calculate percent change varibles
us.trcts.acs.chng <- us.trcts.acs.wide %>% mutate(
  PopChng = ((v16.TotPop - v15.TotPop)/v15.TotPop),
  PovChng = ((v16.BelowPov - v15.BelowPov)/v15.BelowPov),
  MdnIncChng = ((v16.MdnFamInc - v15.MdnFamInc)/v15.MdnFamInc),
  UnemplChng = (((v16.Unempl/v16.CivLabor) - (v15.Unempl/v15.CivLabor))/(v15.Unempl/v15.CivLabor)),
  VacChng = (((v16.Vacant/v16.HousingUnits) - (v15.Vacant/v15.HousingUnits))/(v15.Vacant/v15.HousingUnits)),
  RentChng = (((v16.OccRental/v16.OccUnits) - (v15.OccRental/v15.OccUnits))/(v15.OccRental/v15.OccUnits)),
  GrEduChng = (((v16.MastDegree + v16.DoctDegree)/v16.Pop25) - ((v15.MastDegree + v15.DoctDegree)/v15.Pop25))/((v15.MastDegree + v15.DoctDegree)/v15.Pop25),
  ColEduChng = ((v16.BachDegree/v16.Pop25) - (v15.BachDegree/v15.Pop25))/(v15.BachDegree/v15.Pop25),
  RaceChng = ((v16.White/v16.Black) - (v15.White/v15.Black))/(v15.White/v15.Black)
    ) %>% select(1,50:58)

# calculate single year rate varibles
us.trcts.acs.rt16 <- us.trcts.acs.wide %>% mutate(
  Pop = v16.TotPop,
  PovRt = v16.BelowPov/v16.PovPopDenominator,
  MdnInc = v16.MdnFamInc,
  UnemplRt = v16.Unempl/v16.CivLabor,
  VacRt = v16.Vacant/v16.HousingUnits,
  RentRt = v16.OccRental/v16.OccUnits,
  GrEduRt = (v16.MastDegree + v16.DoctDegree)/v16.Pop25,
  ColEduRt = v16.BachDegree/v16.Pop25,
  PctWhite = v16.White/v16.TotPop
    ) %>% select(1,50:58)
#----------------------------------

# calculate single year rate varibles
frmt.vars <- function(x){
  x %>% mutate(
    Pop = v16.TotPop,
    PovRt = v16.BelowPov/v16.PovPopDenominator,
    MdnInc = v16.MdnFamInc,
    UnemplRt = v16.Unempl/v16.CivLabor,
    VacRt = v16.Vacant/v16.HousingUnits,
    RentRt = v16.OccRental/v16.OccUnits,
    GrEduRt = (v16.MastDegree + v16.DoctDegree)/v16.Pop25,
    ColEduRt = v16.BachDegree/v16.Pop25,
    PctWhite = v16.White/v16.TotPop) %>% 
    select(1,(ncol(.)-8):ncol(.))
}

# run formating for tracts, county, state dataframes
us.trcts.acs.rt16 <- us.trcts.acs16.wide %>% frmt.vars()
us.cnty.acs.rt16 <- us.cnty.acs16.wide %>% frmt.vars()
us.st.acs.rt16 <- us.st.acs16.wide %>% frmt.vars()

```

### 3.Merge attribute data with spatial data
```{r 3 merge attribute data, cache=T}

# merge (left join) 'Opportunity Zones' to sf tracts dataframe 
us.oz.sf <- merge(ustracts_sf.48.albers, ozones.51, by.x = "GEOID", by.y = "GEOID10", all.x)

# merge (left join) NMTC data and ACS attributes to Ozone tracts sf df
us.oz.sf2 <- us.oz.sf %>% 
  select(-c(variable,estimate,moe)) %>% 
  merge(nmtc.tract, by="GEOID", all.x=T) %>% 
  merge(us.trcts.acs.chng,by="GEOID", all.x=T) #use percent change ACS variables

us.oz.sf3 <- us.oz.sf %>% 
  select(-c(variable,estimate,moe)) %>% 
  merge(nmtc.tract, by="GEOID", all.x=T) %>% 
  merge(us.trcts.acs.rt16,by="GEOID", all.x=T) %>% #use 2016 rates ACS varaibles
  mutate(CountyFIPS=substr(GEOID,1,5)) %>% 
  merge(us.cnty.acs.rt16[,c("GEOID","PovRt","MdnInc")], by.x="CountyFIPS",by.y="GEOID", all.x=T) %>% 
  rename(PovRt.trct = PovRt.x) %>% 
  rename(MdnInc.trct = MdnInc.x) %>% 
  rename(PovRt.cnty = PovRt.y) %>% 
  rename(MdnInc.cnty = MdnInc.y) %>% 
  mutate(StateFIPS=substr(GEOID,1,2)) %>% 
  merge(us.st.acs.rt16[,c("GEOID","PovRt","MdnInc")], by.x="StateFIPS",by.y="GEOID", all.x=T) %>% 
  rename(PovRt.st = PovRt) %>% 
  rename(MdnInc.st = MdnInc) %>% 
  mutate(PovRatio.cnty = PovRt.trct/PovRt.cnty) %>% 
  mutate(PovRatio.st = PovRt.trct/PovRt.st) %>% 
  mutate(MdnIncRatio.cnty = MdnInc.trct/MdnInc.cnty) %>% 
  mutate(MdnIncRatio.st = MdnInc.trct/MdnInc.st)

```

### 4.Descriptive analysis
#### 4a.National and State Distribution
```{r 4a national and state descriptive tables/charts, cache=T}

# two-way table of eligibility and designation
table(us.oz.sf3$QOZ,us.oz.sf3$TYPE)
prop.table(table(us.oz.sf3$QOZ,us.oz.sf3$TYPE))

# bar chart: state level distribution of tract deisngation/eligibility
us.oz.sf3 %>% 
  st_set_geometry(NULL) %>% 
  mutate(OZ = factor(ifelse(QOZ=="DESIGNATEDQOZ", "OZone", ifelse(TYPE=="LIC", "Non-OZ LIC", "Non-LIC")))) %>% 
  group_by(state) %>% 
  summarise(n = n(), OZone = sum(OZ == "OZone")/n,LIC = sum(OZ == "Non-OZ LIC")/n,NonLIC = sum(OZ == "Non-LIC")/n) %>% 
  mutate(state = factor(state, levels = state[order(OZone)])) %>% 
  gather(key = "OZdesignation", value = "n.tracts", OZone, LIC, NonLIC) %>% 
  ggplot(aes(x = state, y = n.tracts, fill = factor(OZdesignation, levels = c("NonLIC","LIC","OZone")))) + 
  theme_bw() + geom_bar(stat = "identity") + 
  ylab("% of state tracts") + 
  labs(fill = "Census Tract\nEligibility/Designation") + 
  scale_fill_manual(values=c("grey","blue","dark blue"))

```                     

#### 4b.Compare OZ and Non-OZ-LIC pt1
```{r 4b Compare OZ and Non-OZ-LIC pt1, cache=T}

# select dataset, and remove geometry (seems to increase processing speed), add area/density variables
d <- us.oz.sf3 %>% 
  mutate(area = set_units(st_area(.),mi^2)) %>% 
  st_set_geometry(NULL) %>% 
  filter(TYPE != "NO") %>% #restrict to LIC (low-income communities == QOZ eligible)
  mutate(QOZ = as.factor(QOZ)) %>%  #set QOZ as factor 
  mutate(popdens = as.double(Pop/area)) %>% 
  mutate(prjdens = as.double(n.prj/area)) %>% 
  mutate(invdens = as.double(ttl.amt.prj/area))

#table one, compare all variables between OZ and Non-OZ-LIC
tableOne1 <- CreateTableOne(strata = "QOZ", vars = c("Pop","popdens","PovRt.trct","MdnInc.trct","UnemplRt","VacRt","RentRt","GrEduRt","ColEduRt","PctWhite"), data = d)
tableOne1

tableOne2 <- CreateTableOne(strata = "QOZ", vars = c("metro","n.cde","n.prj","avg.prj.yr","ttl.amt.prj","avg.amt.prj","avg.amt.yr","CDE","NRE","RE","SPE"), data = d)
tableOne2

tableOne3 <- CreateTableOne(strata = "QOZ", vars = c("PovRatio.cnty","PovRatio.st","MdnIncRatio.cnty","MdnIncRatio.st","prjdens","invdens"), data = d)
tableOne3

# compare distribution of Poverty Rate between OZ and Non-OZ LIC
d %>% ggplot(aes(PovRt.trct, color = QOZ)) + 
  scale_color_manual(values=c("#E69F00", "#56B4E9"), 
                       name="Census Tract\nDesignation",
                       breaks=c("DESIGNATEDQOZ","NO"),
                       labels=c("Designated OZone", "Non-OZone LIC")) +
  geom_density(alpha = 0.4) + 
  geom_vline(data=data.frame(d[!is.na(d$PovRt.trct),] %>% 
                               group_by(QOZ) %>% 
                               summarise(grp.mean=mean(PovRt.trct))), 
    aes(xintercept=grp.mean, color=QOZ), 
    linetype="dashed") +
    xlab("2016 Poverty Rate") 
  
```                     

#### 4c.Compare OZ and Non-OZ-LIC pt2
```{r 4c Compare OZ and Non-OZ-LIC pt2, cache=T}

# review distribution of variables
skim(d[23:39])

#cont var x bin/cat >> box plots

p1 <- ggboxplot(d, x = "QOZ", y = "PovRt.trct",
                 color = "QOZ", palette = "jco")
p2 <- ggboxplot(d, x = "QOZ", y = "UnemplRt",
                 color = "QOZ", palette = "jco")
p3 <- ggboxplot(d, x = "QOZ", y = "VacRt",
                 color = "QOZ", palette = "jco")
p4 <- ggboxplot(d, x = "QOZ", y = "GrEduRt",
                 color = "QOZ", palette = "jco")
p5 <- ggboxplot(d, x = "QOZ", y = "ColEduRt",
                 color = "QOZ", palette = "jco")
p6 <- ggboxplot(d, x = "QOZ", y = "PctWhite",
                 color = "QOZ", palette = "jco")

ggarrange(p1,p2,p3,p4,p5,p6 + rremove("x.text"), 
          labels = c("A", "B", "C", "D", "E", "F"),
          ncol = 3, nrow = 2)

```                     

### 5.Map data
```{r 5 generate maps, cache=T, eval=T}

states.sf <- usstates_sf.48.albers %>% 
  merge(data.frame(unique(fips_codes[,1:2])),by.x="GEOID",by.y="state_code") %>% 
  merge(stab,by="state")

counties.sf <- uscounties_sf.48.albers %>% 
  merge(data.frame(unique(fips_codes[,1:2])),by.x="GEOID",by.y="state_code") %>% 
  merge(stab,by="state")

mkmaps <- function(x,y,z,w){
x %>%
  filter(state %in% z) %>% 
  group_by(eval(parse(text=w))) %>% 
  summarise(value = first(eval(parse(text=y)))) %>% 
  ggplot(aes(fill = value)) +
  geom_sf(color = NA, lwd=.5) +
  labs(fill=y) +
  geom_sf(data = st_geometry(states.sf[states.sf$state %in% z,]), fill = "transparent", color = "white", lwd=.5) +
  geom_sf(data = st_geometry(x[(x$state %in% z) & x$QOZ=="DESIGNATEDQOZ",]), fill = "transparent", color = "orange", lwd=.5) %>%
    return()
}

#eval(parse(text=y))

s <- us.oz.sf3 %>% 
  mutate(area = set_units(st_area(.),mi^2)) %>% 
  mutate(QOZ = as.factor(QOZ)) %>%  #set QOZ as factor 
  mutate(popdens = as.double(Pop/area)) %>% 
  mutate(prjdens = as.double(n.prj/area)) %>% 
  mutate(invdens = as.double(ttl.amt.prj/area))

tract.qoz.al <- s %>% mkmaps('TYPE',c("AL"),"GEOID")
tract.qoz.pa <- s %>% mkmaps('TYPE',c("PA"),"GEOID")
tract.qoz.ny <- s %>% mkmaps('TYPE',c("NY"),"GEOID")

cowplot::plot_grid(
  tract.qoz.al,
  tract.qoz.pa,
  tract.qoz.ny,
  nrow = 1
)

tract.povrt.map <- s %>% mkmaps('PovRt.trct',c("AL"),'GEOID')
county.povrt.map <- s %>% mkmaps('PovRt.cnty',c("AL"),'CountyFIPS')
state.povrt.map <- s %>% mkmaps('PovRt.st',c("AL","MS"),'StateFIPS')

cowplot::plot_grid(
  tract.povrt.map,
  county.povrt.map,
  state.povrt.map,
  nrow = 1
)



```


### 6.Spatial Analysis
```{r, eval=FALSE}
#spatial patterns

# ANALYSIS QUESTION -- where OZones more or less likely than non-designated, but eligible LICs to be clustered among other tracts of similar neighborhood characteristics
#- distress index?
#- poverty?

# ANALYSIS QUESTION -- describe spatial patterns of OZones: are different states more or less clustered??
require("spdep")
require("magrittr")
require("map")

# initialize global moran's I function
globalSA <- function(x,y){
  d <- as(x[!is.na(st_dimension(x)),], 'Spatial')
  #set up weight matrix
  w <- poly2nb(d, row.names=d$NAMES)
  wm <- nb2mat(w, style='B')
  
  #row standardisation of the weights matrix:
  rwm <- mat2listw(wm, style='W')
  
  # Checking if rows add up to 1
  mat <- listw2mat(rwm)
  apply(mat, 1, sum)[1:15]
  vec <-as.numeric(d[y][[1]])
  p1 <- moran.plot(vec, rwm)

  #compute Moran’s I and do a significance test
  ww <-  nb2listw(w, style='B')

  #test for significance. 
  sig <- moran.test(vec, ww, randomisation=FALSE)
  
  #Monte Carlo simulation
  mc <- moran.mc(vec, ww, nsim=99)
  
  return(list(vec, rwm, sig,mc))
}

morans_model <- function(df) {
  globalSA(df,"OZone")
}

by_state <- 
  us.oz.sf %>% 
  #filter(state %in% c("AL")) %>% 
  filter(state_code %in% lower48[27:51]) %>% 
  mutate(OZone = (QOZ == "DESIGNATEDQOZ")) %>% 
  group_by(state) %>% 
  nest()

by_state <- by_state %>% 
  mutate(morans_model = purrr::map(data, morans_model)) 

get_morans.i <- function(df) {
  return(df[4][[1]]$statistic)
}

get_p.val <- function(df) {
  return(df[4][[1]]$p.value)
}

by_state <- by_state %>% 
  mutate(morans_i = purrr::map(morans_model, get_morans.i)) %>% 
  mutate(morans_pval = purrr::map(morans_model, get_p.val))

```

### 7.Temporal Analysis
```{r}
#time trends

```

### 8.Regression Analysis
```{r}
#outcomes... classification as QOZ
#2-sample tests, compare LIC/QOZ
chisq.test(table(d$QOZ, d$PovRt.trct)) 
summary((glm(QOZ~Pop + PovRt.trct + MdnInc.trct + UnemplRt + VacRt + RentRt + ColEduRt + GrEduRt + PctWhite, data=d, family=binomial())))
summary((glm(QOZ~n.cde + n.prj + avg.prj.yr + ttl.amt.prj + avg.amt.prj + avg.amt.yr, data=d, family=binomial())))
exp(coef(glm(QOZ~PovRt.trct, data=d, family=binomial()))) #Odds ratios

```



### 9.Web-map, leaflet, Shiny tool
```{r}
# make a leaflet map
# make a shiny tool to select and display state maps

```


