---
title: "finalproject_dnkarp"
author: "dnkarp"
date: "10/26/2018"
output: html_document
---

#--------------------------------------------------------------
"Social impact could include, but not limited to; workforce education/development, workforce diversity, neighborhood revitalization, health and environment friendly and job creation." - WNY Impact Investment Fund

RULE: a low-income community census tract has an individual poverty rate of at least 20% and median family income up to 80% percent of the area median. Every state or territory can designate up to 25% of its census tracts that meet qualification requirements

EXCEPTION: 5% of census tracts that do not meet the definition of a low-income community can be designated under an exemption. Exempt census tracts must be contiguous with low-income community census tracts that are designated as Opportunity Zones, and the median family income of the exempt tract must not exceed 125% of the median family income of the designated low-income community census tract with which it is contiguous. 

DEFINITIONS: low-income community 45d(e) [American Jobs Creation Act of 2004, IRC §45D(e)(2)]
A “low-income community” is defined as any population census tract where the 
- poverty rate for such tract is at least 20% 
- or, in the case of a tract not located within a metropolitan area, median family income for such tract does not exceed 80 of statewide median family income, 
- or, in the case of a tract located within a metropolitan area, the median family income for such tract does not exceed 80% of the greater of statewide median family income or the metropolitan area median family income.

--------------------------------------------------------------
exploring social impact investments -- how many CDFIs, CDEs, CDCs are opporating in the 'region'... NMTCs?
number of impact funds?
indicators of:
  workforce education/development, 
  workforce diversity, 
  neighborhood revitalization, 
  health and environment friendly,
  and job creation.
  
bond issuance?

#--------------------------------------------------------------
### 0.Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### 1.Load libraries
```{r load libs, message=F, warning=F, cache=T, results='hide'}

# data wrangling libs
library(tidyverse)
library(broom)
library(data.table)
library(purrr)
library(stringr)

# data analysis
library(skimr)
library(units)

# spatial libs
library(sf)
library(maps)
library(albersusa)

# census data libs + install api key
library(tidycensus)
library(tigris)
census_api_key("e7051ec917473383ba64a0c65d7090fa5cc10cf4")
options(tigris_use_cache = TRUE)

# graphics/viz libs
library(viridis)
library(gridExtra)
library(ggthemes)
library(cowplot)


```

### 2.Load data
##### 2a.Load data
```{r 2a load OZ csv data, cache=T}

# load opportunity zones (source: CDFI Fund[https://www.cdfifund.gov/Pages/Opportunity-Zones.aspx]); csv, tract-level data (all US)
ozones.raw <- read_csv("data/OpportunityZones2/Opportunity Zones.csv")

# get list of state postal codes, assign state/territory flag
# 50+DC; territories (PR - Puerto Rico, VI - US Vir Islands, GU - Guam, MP - N.Mariana, UM - Minor Outlying, AS - Am Samoa)
stab <- rbind(
   cbind(state=c(unique(fips_codes$state)[1:51]),is.state=as.logical(T))
  ,cbind(state=c(unique(fips_codes$state)[52:57]),is.state=as.logical(F))
)

# merge (left join) state abbreviations to ozone data (based on FIPS code); add state/terr flag
ozones <- ozones.raw %>% mutate("state_code"=substr(ozones.raw$GEOID10,1,2)) %>% merge(data.frame(unique(fips_codes[,1:2])),by="state_code") %>% merge(stab,by="state")

# set as datatable
setDT(ozones)

# subset ozone data if in 51 or in territories
ozones.51 <- ozones[as.logical(is.state),]
ozones.terr <- ozones[!as.logical(is.state),]

```

##### 2b.Load data
```{r 2b load ACS spatial features data from raw, cache=T}
if(!file.exists("data/uscounties_sf.rds")){
# download census tract boundaries (with population data, 2016) from tidycensus; get per state and bind
ustracts_sf <- reduce(
  purrr::map(stab[stab[,2]==T,1], function(x) {
    get_acs(geography = "tract", variables = "B01003_001", 
            state = x, geometry = TRUE)
  }), 
  rbind
)

# get counties (filter PR)
uscounties_sf <- get_acs(geography = "county", variables = "B01003_001", geometry = TRUE) %>% filter(substr(GEOID,1,2)!="72")
# get states (filter PR)
usstates_sf <- get_acs(geography = "state", variables = "B01003_001", geometry = TRUE) %>% filter(GEOID!="72")

# Save a sf object to a file
saveRDS(ustracts_sf, "data/ustracts_sf.rds")
saveRDS(uscounties_sf, "data/uscounties_sf.rds")
saveRDS(usstates_sf, "data/usstates_sf.rds")
}
```

##### 2c.Load data
```{r 2c load ACS spatial features data from repo, cache=T}
# Restore it under a different name
ustracts_sf <- readRDS("data/ustracts_sf.rds")
uscounties_sf <- readRDS("data/uscounties_sf.rds")
usstates_sf <- readRDS("data/usstates_sf.rds")

# reproject to albers & drop hawaii and alaska (for now)
lower48 <- data.frame(state_code=unique(fips_codes$state_code)) %>% filter(!(state_code %in% c("02","15"))) %>% mutate(state_code=as.character(state_code)) %>% .[[1]]

prj.alb.48 <- function(x){x %>% filter(substr(GEOID,1,2) %in% lower48) %>% st_transform(us_laea_proj)}

ustracts_sf.48.albers <- prj.alb.48(ustracts_sf)
uscounties_sf.48.albers <- prj.alb.48(uscounties_sf)
usstates_sf.48.albers <- prj.alb.48(usstates_sf)

```

##### 2d.Load data
```{r 2d load NMTC csv data from raw, cache=T}
nmtc.raw <- read_csv("data/nmtc_investments.csv")
#NMTC Public Data Release: 2003-2015 Data File [https://www.cdfifund.gov/news-events/news/Pages/news-detail.aspx?NewsID=270&Category=Press%20Releases]

nmtc <- nmtc.raw %>% 
  mutate(GEOID = str_pad(`2010 Census Tract`, width=11, side="left", pad="0")) %>% 
  mutate(metro = as.factor(`Metro/Non-Metro, 2010 Census`)) %>% 
  mutate(cde = as.factor(`Community Development Entity (CDE) Name`)) %>% 
  mutate(type = as.factor(`QALICB Type`)) %>% 
  mutate(year = `Origination Year`) %>% 
  mutate(qlici.amount = `Project QLICI Amount`) %>% 
  mutate(total.cost = `Estimated Total Project Cost`) %>% 
  select(GEOID, metro, cde, type, year, qlici.amount, total.cost)

nmtc.tract <- nmtc %>%
  group_by(GEOID, type) %>% 
  mutate(one = 1) %>% 
  summarize(one = sum(one)) %>% 
  spread(type, one) %>% 
  group_by(GEOID) %>% 
  summarise_all(funs(sum))

nmtc.tract <- nmtc %>% 
  group_by(GEOID) %>% 
  summarize(metro = first(metro), 
            n.cde = length(unique(cde)), 
            n.prj = n(), 
            avg.prj.yr = n()/length(unique(year)), 
            ttl.amt.prj = sum(qlici.amount), 
            avg.amt.prj = mean(qlici.amount), 
            avg.amt.yr = sum(qlici.amount)/length(unique(year))) %>%
  merge(nmtc.tract, by="GEOID")

```

### 3.Map data
```{r 3 make 'test' maps, eval=F}
mkmaps <- function(x,y){
x %>%
  ggplot(aes(fill = eval(parse(text=y)))) + 
  geom_sf(color = NA, lwd=.5) +
  labs(fill=y) %>% 
    return()
}

tract.pop.map <- ustracts_sf.48.albers %>% filter(substr(GEOID,1,2)=="01") %>% mkmaps('estimate')
county.pop.map <- uscounties_sf.48.albers%>% filter(substr(GEOID,1,2)=="01") %>% mkmaps('estimate')
state.pop.map <- usstates_sf.48.albers %>% filter(substr(GEOID,1,2)=="01") %>% mkmaps('estimate')

cowplot::plot_grid(
  tract.pop.map,
  county.pop.map,
  state.pop.map,
  nrow = 1
)

```

### 4.Merge attribute data with spatial data
#### 4a.Merge OZone attribute data with spatial data
```{r 4 merge attribute data, cache=T}
us.oz.sf <- merge(ustracts_sf.48.albers, ozones.51, by.x = "GEOID", by.y = "GEOID10", all.x)

if(F){

tract.qoz.al <- us.oz.sf %>% filter(substr(GEOID,1,2)=="01") %>% mkmaps('QOZ')
tract.qoz.pa <- us.oz.sf %>% filter(substr(GEOID,1,2)=="42") %>% mkmaps('QOZ')
tract.qoz.ny <- us.oz.sf %>% filter(substr(GEOID,1,2)=="36") %>% mkmaps('QOZ')


cowplot::plot_grid(
  tract.qoz.al,
  tract.qoz.pa,
  tract.qoz.ny,
  nrow = 1
)

}

```

#### 4b.Get and Merge area(community) variables
```{r}
#poverty
#race/ethnic mix
#education, employment, housing, health
#commerical development, jobs, industry, GDP??
#environment: urban/suburban/rural/metro-region

# load variables
v16 <- load_variables(2016, "acs5", cache = TRUE)
v16 %>% select(name, label) %>% filter(grepl("19113", name))

# population
totpop_var <- cbind(c("B00001_001"),c("TotPop"))
# poverty
# https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&var=B17001002
# https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&var=B19113001
povinc_var <- cbind(c("B17001_002","B19113_001"),c("BelowPov","MdnFamInc"))
# race/ethnic mix
# https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001
race_var <- cbind(
  c("B02001_002","B02001_003","B02001_004","B02001_005","B02001_006"),
  c("White","Black","NatAm","Asian","NatIsl")
  )
#education, employment, housing
# https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B15003
# https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B23025
# https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B25002
# https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B25003
socioecon_var <- cbind(
  c("B15003_001", "B15003_017", "B15003_018", "B15003_019","B15003_020", "B15003_021","B15003_022","B15003_023","B15003_024", "B15003_025", "B23025_003", "B23025_005", "B25002_001", "B25002_003", "B25003_001", "B25003_003"),
  c("Pop25", "HS", "GED", "CollegeNoDegree0", "CollegeNoDegree1", "AssocDegree", "BachDegree", "MastDegree", "ProfDegree", "DoctDegree", "CivLabor", "Unempl", "HousingUnits", "Vacant", "OccUnits", "OccRental" )
  )
#commerical development, jobs, industry, GDP??
econdev_var <- c()
#urban: urban/suburban/rural/metro-region
urban_var <- c()
#health: diabetes, obesity, substance abuse, chronic
#buit environment: parks, rec centers, health clinics, civic inst?
#property value?

select_vars <- rbind(totpop_var, povinc_var, race_var, socioecon_var, econdev_var, urban_var) %>% data.frame() %>% rename(variable = X1, v = X2) %>% mutate(variable = as.character(variable))

us <- unique(fips_codes$state)[1:51]

us.trcts.acs16 <- reduce(
  purrr::map(us[1], function(x) {
    get_acs(geography = "tract", year = 2016, variables = select_vars[,1], 
            state = x, geometry = F)
  }), 
  rbind
)

us.trcts.acs15 <- reduce(
  purrr::map(us[1], function(x) {
    get_acs(geography = "tract", year = 2015, variables = select_vars[,1], 
            state = x, geometry = F)
  }), 
  rbind
)


us.trcts.acs16.wide <- us.trcts.acs16 %>% left_join(select_vars, by="variable") %>% select(-c(moe,NAME,variable)) %>% 
spread(key = v, value = estimate, sep = "16.")

us.trcts.acs15.wide <- us.trcts.acs15 %>% left_join(select_vars, by="variable") %>% select(-c(moe,NAME,variable)) %>% 
spread(key = v, value = estimate, sep = "15.")

us.trcts.acs.wide <- merge(us.trcts.acs16.wide,us.trcts.acs15.wide,by="GEOID")

us.trcts.acs.chng <- us.trcts.acs.wide %>% mutate(
  PopChng = ((v16.TotPop - v15.TotPop)/v15.TotPop),
  PovChng = ((v16.BelowPov - v15.BelowPov)/v15.BelowPov),
  MdnIncChng = ((v16.MdnFamInc - v15.MdnFamInc)/v15.MdnFamInc),
  UnemplChng = (((v16.Unempl/v16.CivLabor) - (v15.Unempl/v15.CivLabor))/(v15.Unempl/v15.CivLabor)),
  VacChng = (((v16.Vacant/v16.HousingUnits) - (v15.Vacant/v15.HousingUnits))/(v15.Vacant/v15.HousingUnits)),
  RentChng = (((v16.OccRental/v16.OccUnits) - (v15.OccRental/v15.OccUnits))/(v15.OccRental/v15.OccUnits)),
  GrEduChng = (((v16.MastDegree + v16.DoctDegree)/v16.Pop25) - ((v15.MastDegree + v15.DoctDegree)/v15.Pop25))/((v15.MastDegree + v15.DoctDegree)/v15.Pop25),
  ColEduChng = ((v16.BachDegree/v16.Pop25) - (v15.BachDegree/v15.Pop25))/(v15.BachDegree/v15.Pop25),
  RaceChng = ((v16.White/v16.Black) - (v15.White/v15.Black))/(v15.White/v15.Black)
    ) %>% select(1,50:58)

us.trcts.acs.rt16 <- us.trcts.acs.wide %>% mutate(
  Pop = v16.TotPop,
  PovRt = v16.BelowPov/v16.TotPop,
  MdnInc = v16.MdnFamInc,
  UnemplRt = v16.Unempl/v16.CivLabor,
  VacRt = v16.Vacant/v16.HousingUnits,
  RentRt = v16.OccRental/v16.OccUnits,
  GrEduRt = (v16.MastDegree + v16.DoctDegree)/v16.Pop25,
  ColEduRt = v16.BachDegree/v16.Pop25,
  PctWhite = v16.White/v16.TotPop
    ) %>% select(1,50:58)

us.oz.sf2 <- us.oz.sf %>% select(-c(variable,estimate,moe)) %>% merge(nmtc.tract, by="GEOID", all.x=T) %>% merge(us.trcts.acs.chng,by="GEOID", all.x=T)

us.oz.sf3 <- us.oz.sf %>% select(-c(variable,estimate,moe)) %>% merge(nmtc.tract, by="GEOID", all.x=T) %>% merge(us.trcts.acs.rt16,by="GEOID", all.x=T)

```

### 5.Descriptive analysis
```{r}
#national
#regional
#state

d <- us.oz.sf3 %>% st_set_geometry(NULL)

#define batch plotting functions
batchbox <- function(x,y){
  ggplot(d, aes(x=x, y=y)) +
    geom_boxplot(color="blue")
  }

batchbox("QOZ","PopChng")

d %>% filter(TYPE == "LIC") %>% 
  ggplot(aes(x=factor(QOZ), y=UnemplRt)) +
    geom_boxplot()

my_cols <- c("#00AFBB", "#E7B800", "#FC4E07")
pairs(d[(d$TYPE != "NO"),11:16], pch = 19,  cex = 0.5,
      col = my_cols[as.factor(d$QOZ)],
      lower.panel=NULL)

# get crosstab of number of tracts by LIC and QOZ designations
d %>% group_by(TYPE, QOZ) %>% xtabs(~ TYPE + QOZ, data=.)
s <- d %>% group_by(state) %>% summarize(n_tract = n(), n_LIC = sum(TYPE == "LIC"), pct_LIC = mean(TYPE == "LIC"), n_QOZ = sum(QOZ == "DESIGNATEDQOZ"), pct_QOZ = mean(QOZ == "DESIGNATEDQOZ"), pct_QOZ2 = n_QOZ/n_LIC) 

# get top and bottom deciles of % of tracts designated as LICs
s %>% arrange(pct_LIC) %>% slice(c(1:5,(n()-4):n()))
# get top and bottom deciles of % of LICs designated as QOZ
s %>% arrange(pct_QOZ2) %>% slice(c(1:5,(n()-4):n()))

#total area of QOZ, by state
area_by_state <- us.oz.sf %>% 
  mutate(area = st_area(.)) %>% 
  st_set_geometry(NULL) %>% 
  group_by(state, QOZ) %>% 
  summarize(area = set_units(sum(area),km^2)) 

# plot distribution of QOZ by state
d %>% filter(QOZ != "NO") %>% group_by(state) %>% ggplot(., aes(x=state)) + 
  geom_bar(aes(y = (..count..)),fill="orange", alpha=0.6) + 
  geom_text(aes(y = (..count..) ,label =   ifelse((..count..)==0,"",scales::percent((..count..)/sum(..count..)))), stat="count",colour="darkgreen") + 
  theme_bw() +
  xlab("Designation") +
  ylab("Frequency") + 
  ggtitle("Distribution of Designated Tracts by State") +
  coord_flip()

# plot stacked QOZ by state (how to reoder)
d %>% mutate(QOZ = as.factor(QOZ)) %>% group_by(state) %>% summarize(frq =  mean(as.numeric(QOZ))) %>% mutate(x = forcats::fct_reorder(state, frq)) %>% 
  ggplot(aes(x, fill = frq))+
  geom_bar(position="fill") +
  scale_y_continuous(labels = percent)

data %>% 
  mutate(x = forcats::fct_reorder(x, as.numeric(z), fun = mean)) %>% 
  ggplot(aes(x, fill = z)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent)

d %>% group_by() %>% arrange() %>% mutate(QOZ = as.factor(QOZ)) %>% ggplot(aes(state, fill=QOZ)) +
    geom_bar(position="fill")

```

### 6.Spatial Analysis
```{r}
#spatial patterns

# ANALYSIS QUESTION -- where OZones more or less likely than non-designated, but eligible LICs to be clustered among other tracts of similar neighborhood characteristics
- distress index?
- poverty?

# ANALYSIS QUESTION -- describe spatial patterns of OZones: are different states more or less clustered??
require("spdep")
require("magrittr")
require("map")

# initialize global moran's I function
globalSA <- function(x,y){
  d <- as(x[!is.na(st_dimension(x)),], 'Spatial')
  #set up weight matrix
  w <- poly2nb(d, row.names=d$NAMES)
  wm <- nb2mat(w, style='B')
  
  #row standardisation of the weights matrix:
  rwm <- mat2listw(wm, style='W')
  
  # Checking if rows add up to 1
  mat <- listw2mat(rwm)
  apply(mat, 1, sum)[1:15]
  vec <-as.numeric(d[y][[1]])
  p1 <- moran.plot(vec, rwm)

  #compute Moran’s I and do a significance test
  ww <-  nb2listw(w, style='B')

  #test for significance. 
  sig <- moran.test(vec, ww, randomisation=FALSE)
  
  #Monte Carlo simulation
  mc <- moran.mc(vec, ww, nsim=99)
  
  return(list(vec, rwm, sig,mc))
}

morans_model <- function(df) {
  globalSA(df,"OZone")
}

by_state <- 
  us.oz.sf %>% 
  #filter(state %in% c("AL")) %>% 
  filter(state_code %in% lower48[27:51]) %>% 
  mutate(OZone = (QOZ == "DESIGNATEDQOZ")) %>% 
  group_by(state) %>% 
  nest()

by_state <- by_state %>% 
  mutate(morans_model = purrr::map(data, morans_model)) 

get_morans.i <- function(df) {
  return(df[4][[1]]$statistic)
}

get_p.val <- function(df) {
  return(df[4][[1]]$p.value)
}

by_state <- by_state %>% 
  mutate(morans_i = purrr::map(morans_model, get_morans.i)) %>% 
  mutate(morans_pval = purrr::map(morans_model, get_p.val))

```

### 7.Temporal Analysis
```{r}
#time trends

```

### 8.Regression Analysis
```{r}
#outcomes... classification as QOZ
#2-sample tests, compare LIC/QOZ

```


