---
title: "geo503-2018-finalproject-dnkarp"
author: "dnkarp"
date: "October 25, 2018"
output: html_document
---

### 0.Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### 1.Load libraries
```{r load libs, message=F, warning=F, results='hide'}
library(tidyverse)
library(sf)
library(maps)
library(stringr)
library(tidycensus)
library(viridis)
library(tigris)
library(skimr)
library(gridExtra)
library(ggthemes)
library(purrr)
library(broom)
library(data.table)

```

### 2.Load data
```{r 2a}
ozones2 <- st_read("data/OpportunityZones2/shp/theme2.shp")
ozones.raw <- read_csv("../data/OpportunityZones2/Opportunity Zones.csv")
setDT(ozones.raw)

unique(fips_codes[,2:3])
stcodes.terr <- c("60","66","69","72","74","78")

unique(fips_codes[,1:2])
ozones <- ozones.raw %>% mutate("state_code"=substr(ozones.raw$GEOID10,1,2)) %>% merge(data.frame(unique(fips_codes[,1:2])),by="state_code")
stcodes.51 <- unique(fips_codes$state_code[!(fips_codes$state_code %in% stcodes.terr)])

ozones.51 <- ozones[substr(ozones$GEOID10,1,2) %in% stcodes.51,]
ozones.terr <- ozones[substr(ozones$GEOID10,1,2) %in% stcodes.terr,]

unique(ozones.51$STATENAME)
unique(ozones.51$state_code)

statelist <- unique(fips_codes[fips_codes$state_code %in% stcodes.51,"state"])

```

### 3.Select variables
```{r 3a}
# look at raw data import
names(data.import)
str(data.import)

# count obs with no ZIP code
data.import %>% filter(Zip.Code == "0") %>% count()

# select, rename variables, and drop obs w/ missing ZIP codes
dt.select <- data.import %>% 
  filter(Zip.Code != "0") %>% 
  select(
    Year, 
    City, 
    County, 
    Zip.Code,
    People=..People,
    Elderly=..Elderly,
    Age.of.Home,
    Occupancy.Type,
    Housing.Type,
    Total.Floor.Area,
    Blower.Pre=Blower.Door.Reading..Pre.,
    Blower.Post=Blower.Door.Reading..Post.,
    Percent.Reduction,
    Total.Material.Costs
  )

```

### 4.Format case level dataset
```{r 4a}
# reformat City/County variables; drop obs with negative or '99999' blower values

summary(dt.select$Housing.Type)
summary(dt.select$Occupancy.Type)
summary(dt.select$Age.of.Home)
summary(subset(dt.select,Age.of.Home>100)$Age.of.Home)

skim(dt.select)

dt.select %>% filter(Blower.Post == 99999) %>% count()
dt.select %>% filter(Blower.Pre <= 0) %>% count()
dt.select %>% filter(Blower.Post == 99999) %>% count()
dt.select %>% filter(Blower.Pre <= 0) %>% count()

dt.clean <- dt.select %>% mutate(City.chr = as.character(City)) %>% 
                          mutate(Zip.chr = as.character(Zip.Code)) %>% 
                          mutate(County.chr = as.character(County)) %>% 
                          mutate(County.chr = substr(County.chr,2,nchar(County.chr))) %>% 
                          separate(County.chr, c("County.num","County.nam"),sep = "\\) ") %>% 
                          filter(grepl("^[A-Za-z]+$", substr(City.chr,1,1))) %>% 
                          mutate(Housing.Type = toupper(Housing.Type)) %>%
                          filter(grepl("[[:alpha:]]", Housing.Type)) %>%
                          mutate(Housing.Type = as.factor(Housing.Type)) %>% 
                          mutate(Occupancy.Type = as.factor(Occupancy.Type)) %>% 
                          filter(Blower.Post != 99999) %>% 
                          filter(Blower.Post >= 0) %>% 
                          filter(Blower.Pre != 99999) %>% 
                          filter(Blower.Pre >= 0)
nrow(dt.select)
nrow(dt.clean)

str(dt.clean)

summary(dt.clean$Housing.Type)
summary(dt.clean$Occupancy.Type)

skim(dt.clean)

```

### 5.Format ZIP code level dataset
```{r 5a}
# continguous variables sum median, mean?

# categorical variables count, percent dummies?

```

### 6.Format county level dataset
```{r 6a}
# continguous variables sum median, mean?

# categorical variables count, percent dummies?


```

### 7.Load spatial data
```{r 7a}

# install census api key
key <- "e7051ec917473383ba64a0c65d7090fa5cc10cf4"
census_api_key(key, overwrite = FALSE, install = FALSE)

# load variables
v16 <- load_variables(2016, "acs5", cache = TRUE)
v16 %>% select(name, label) %>% filter(grepl("Total", label))

us <- unique(fips_codes$state)[1:51]

library(sf)
options(tigris_use_cache = TRUE)

totalpop_sf <- reduce(
  map(us, function(x) {
    get_acs(geography = "tract", variables = "B01003_001", 
            state = x, geometry = TRUE)
  }), 
  rbind
)

str(totalpop_sf)

# load US census tracts (with total population data for 2016)
us.tracts <- get_acs(geography = "tract", year = 2016,
                  variables = "B00001_001", state = "PA", geometry = T, keep_geo_vars = T)





# load Ohio counties (with total population for 2013)
ohio.counties <- get_acs(state = "OH", geography = "county", year = 2013,
                  variables = "B00001_001", geometry = TRUE)

# union counties into state boundary
ohio.state <- st_sf(st_union(ohio.counties))

ohio.zipcodes.vector <- st_join(st_centroid(us.zipcodes), ohio.counties) %>% filter(GEOID.y != "NA") %>% select(GEOID.x) %>% st_set_geometry(NULL)

ohio.zipcodes <- us.zipcodes %>% filter(GEOID %in% ohio.zipcodes.vector[[1]])

if(F){
plot(st_geometry(ohio.zipcodes))
plot(st_geometry(ohio.counties))
plot(st_geometry(ohio.state))
plot(st_geometry(st_centroid(ohio.zipcodes)), pch = 1, col = 'red')
plot(ohio.zipcodes["estimate"])


ohio.zipcodes %>%
  ggplot(aes(fill = estimate, color = estimate)) + 
  geom_sf() + 
  scale_fill_viridis(option = "magma") + 
  scale_color_viridis(option = "magma") +
  geom_sf(data = ohio.state, fill = NA, color = "grey30", lwd=1)
}
```

### 8.Merge attribute data with spatial data
```{r 8a}
# merge ZIP code
names(dt.clean)
names(ohio.zipcodes)

skim(dt.clean$Zip.chr)
skim(ohio.zipcodes$GEOID)

dt.ohio.zipcodes <- merge(ohio.zipcodes, dt.clean, by.x = "GEOID", by.y = "Zip.chr")

if(F){
gp <- dt.ohio.zipcodes %>%
  group_by(GEOID) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(fill = n, color = n)) + 
  geom_sf() + 
  scale_fill_viridis(option = "magma") + 
  scale_color_viridis(option = "magma") +
  geom_sf(data = ohio.state, fill = NA, color = "grey30", lwd=1)

gp +
    geom_sf(data = ohio.counties, fill = NA, color = "white", lwd=0.5)
}

# merge county
names(dt.clean)
names(ohio.counties)

skim(dt.clean$County.num)
skim(ohio.counties$GEOID)

ohio.clean <- ohio.counties %>% 
  separate(NAME, c("County.nam"),sep = " ")

dt.ohio.counties <- merge(ohio.clean, dt.clean, by = "County.nam")

if(F){
gp.county <- dt.ohio.counties %>%
  group_by(GEOID) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(fill = n, color = n)) + 
  geom_sf() + 
  scale_fill_viridis(option = "magma") + 
  scale_color_viridis(option = "magma") +
  geom_sf(data = ohio.state, fill = NA, color = "grey30", lwd=1)

gp.county
}

```

### 9.Descriptive statistics
```{r 9a}
# merge ZIP code

# merge county

```

#### 10.Distribution of independent variables (continuous)
```{r 10a}
# Generate boxplots

```

#### 11.Distribution of independent variables (categorical)
```{r 11a}
# Generate bar charts

```

#### 12.Distribution of dependent (response) variable
```{r 12a}
# Generate boxplots

```

### 13.Get state totals
```{r 13a}

```

### 14.Time series analysis
```{r 14a}

```

### 15.Generate neighbors matrix
```{r 15a}

```

### 16.Get area, neighbor totals
```{r 16a}

```

### 17.Map data
```{r 17a}
if(F){
dt.ohio.counties %>%
  group_by(GEOID, Occupancy.Type) %>% 
  ggplot(aes(fill = Blower.Pre, color = Blower.Pre)) +
  facet_wrap(~Occupancy.Type) +
  geom_sf()
}

n = nrow(dt.ohio.counties)
c = sum(dt.ohio.counties$Total.Material.Costs, na.rm = T)
  
settheme <-   
  theme_minimal() +
  theme(axis.ticks = element_blank()
        , axis.text.x = element_blank()
        , axis.text.y = element_blank()
        , plot.title = element_text(size = 8)
#        , legend.key.size = unit(1,"cm")
        )

gp.counties <- dt.ohio.counties %>%
  group_by(GEOID) %>% 
  summarise(
            pct = n()/n
          , Blwr.Pre = mean(Blower.Pre)
          , Blwr.Post = mean(Blower.Post)
          , Pct.Red.Avg = mean(Percent.Reduction)
          , Cost = sum(Total.Material.Costs)
          , Cost.pct = Cost/c
          , Cost.avg = mean(Total.Material.Costs))

gp.counties <- gp %>% mutate(Cost.pct2 = (Cost.pct*10000)) %>% mutate(pct2 = (pct*1000))

gp.zipcodes <- dt.ohio.zipcodes %>%
  group_by(GEOID) %>% 
  summarise(
            pct = n()/n
          , Blwr.Pre = mean(Blower.Pre)
          , Blwr.Post = mean(Blower.Post)
          , Pct.Red.Avg = mean(Percent.Reduction)
          , Cost = sum(Total.Material.Costs)
          , Cost.pct = Cost/c
          , Cost.avg = mean(Total.Material.Costs))

gp.zipcodes <- gp.zipcodes %>% mutate(Cost.pct2 = (Cost.pct*100000)) %>% mutate(pct2 = (pct*10000))

plot.maps <- function(gp) {

p1 <- gp %>% 
  ggplot(aes(fill = pct2)) +
  geom_sf(lwd = 0) +
  scale_fill_viridis("pct2", labels = scales::comma) +
  ggtitle("Percent of statewide recipients") +
  settheme

p2 <- gp %>% 
  ggplot(aes(fill = Blwr.Pre)) +
  geom_sf() +
  geom_sf(lwd = 0) +
  scale_fill_viridis("Blwr.Pre", labels = scales::comma) +
  ggtitle("Avg blower test before") +
  settheme

p3 <- gp %>% 
  ggplot(aes(fill = Blwr.Post)) +
  geom_sf() +
  geom_sf(lwd = 0) +
  scale_fill_viridis("Blwr.Post", labels = scales::comma) +
  ggtitle("Avg blower test after") +
  settheme

p4 <- gp %>% 
  filter(Pct.Red.Avg > 0) %>% 
  ggplot(aes(fill = Pct.Red.Avg)) +
  geom_sf() +
  geom_sf(lwd = 0) +
  scale_fill_viridis("Pct.Red.Avg", labels = scales::comma) +
  ggtitle("Avg % reduction in blower test") +
  settheme

p5 <- gp %>% 
  ggplot(aes(fill = Cost)) +
  geom_sf() +
  geom_sf(lwd = 0) +
  scale_fill_viridis("Cost", labels = scales::comma) +
  ggtitle("Total material expenses") +
  settheme

p6 <- gp %>% 
  ggplot(aes(fill = Cost.pct2)) +
  geom_sf() +
  geom_sf(lwd = 0) +
  scale_fill_viridis("Cost.pct2", labels = scales::comma) +
  ggtitle("Percent of statewide material expenses") +
  settheme

p7 <- gp %>% 
  ggplot(aes(fill = Cost.avg)) +
  geom_sf() +
  geom_sf(lwd = 0) +
  scale_fill_viridis("Cost.avg", labels = scales::comma) +
  ggtitle("Average material expenses") +
  settheme

return(list(p1,p2,p3,p4,p5,p6,p7))

}

plots.counties <- plot.maps(gp.counties)
plots.zipcodes <- plot.maps(gp.zipcodes)

grid.arrange(
  grobs = plots.counties,
  widths = c(1, 1, 1, 1),
  layout_matrix = rbind(c(1, 2, 3, 4),
                        c(5, 6, 7, NA))
)

grid.arrange(
  grobs = plots.zipcodes,
  widths = c(1, 1, 1, 1),
  layout_matrix = rbind(c(1, 2, 3, 4),
                        c(5, 6, 7, NA))
)


```

### 18.Test and plot correlations
```{r 18a}
names(dt.ohio.zipcodes)

dt.ohio.zipcodes %>% 
  filter(Percent.Reduction>0) %>% 
  group_by(GEOID) %>% 
  ggplot(aes(x = Percent.Reduction)) +
  geom_density()

dt.ohio.zipcodes[,c("Blower.Pre","Blower.Post")] %>% 
  gather(x,reading,c("Blower.Pre","Blower.Post")) %>% 
  filter(reading<10000) %>% 
  ggplot(aes(x=reading, fill=x)) + geom_density(alpha=0.25)

dt.ohio.zipcodes[,c("Percent.Reduction","Housing.Type")] %>% 
  filter(Percent.Reduction>0) %>% 
  ggplot(aes(x=Housing.Type, y=Percent.Reduction, fill=Housing.Type)) + geom_boxplot(alpha=0.25)

y1<-dt.ohio.zipcodes$Blower.Pre
y2<-dt.ohio.zipcodes$Blower.Post
t.test(y1,y2,paired=TRUE)

dt.ohio.zipcodes %>% filter(Percent.Reduction>0) %>% ggplot(aes(x=Year, y=Blower.Pre, group=Year)) + geom_boxplot()

```

### 19.Estimate linear model
```{r 19a}

```

### 20.Test for spatial autocorrelation
```{r 20a}
# global

# local

```

### 21.Map clusters / hot-spot analysis
```{r 21a}

```

### 22.Describe eligbility criteria 
```{r 22a}

```

### 23.Get area demographics 
```{r 23a}
# tidycensus

```

### 24.Classify areas 
```{r 24a}
# likely to have high exposure

# likely to receive assistance

# likely to see improvement

# likely to have high consumption (relative to house size?)

# likely to have high energy cost burden (relative to area income?)

```

### X.Save workspace 
```{r 24a}
save.image("energy_poverty_data_working.RData")

```